---
title: "Tesla Supercharger Analysis"
author: "Jim-Lee Wu and Fion Huang"
subtitle: 
output:
  pdf_document:
    number_sections: true
  html_document:
    number_sections: true
  word_document: default
---
```{r}
library(tidyverse)
library(corrplot)
```


# Introduction

There has been an increasing rise in popularity in electronic vehicles (EVs) throughout the past decade for its environmental, economic, and performance benefits. With so many EVs filling the streets, a need for a proper EV charging infrastructure is needed. Just like gasoline cars, EV drivers require charging stations where they can recharge when going long distances. 

## Dataset

The data from this data set was gathered via various many online sources. When deciding on location, we wanted to focus on a state that we were familiar with so we chose California. When collecting data, we chose to only collect data from counties that had a population greater than 500,000. The reason we made this decision was because we wanted to focus more on higher populated counties. Our data set contains 16 counties and will contain 296 observations. 

Counties have a combination of towns, unincorporated cities, and incorporated cities. In our data set, we are ONLY considering incorporated cities. 

As for how the data was collected, the parameters: 

`population`, `land_size`, `medium_income`, and `per_capita_income` were scraped from the [US Census](https://www.census.gov/quickfacts/) (Some cities found through [Census Reporter]())

`registered_voter`,	`democratic`, and	`republican` were scraped from [](https://www.sos.ca.gov/elections/report-registration/ror-odd-year-2019/)

`no_stations`,	`more_than_12`,	`no_v2`,	`no_urban`,	`no_v3`,	`no_v4`, and `total` were scraped from []()

`mile` was self obtained from Google Maps using the "Look Up Distance" feature.

```{r}
tesla <- read.csv("Tesla Dataset.csv")
View(tesla) 
```

## How The Data Was Collected


## Parameters

`city` City

`county` County of City

`population` 2022 Population 

`land_size` 2020 Land area in square miles

`medium_income` medium household income in 2018-2022 (in 2022 dollars)

`per_capita_income` per_capita_income: mean income of every person in 2018-2022 (in 2022 dollars)

`no_stations` the number of Tesla Supercharger stations

`mile` the number of stations that are less than one mile from a highway/freeway

`more_than_12` the number of stations that have more than 12 chargers

`no_v2`the number of V2 chargers

`no_urban` the number of urban chargers

`no_v3` the number of V3 chargers

`no_v4` the number of V4 chargers

`total` total number of chargers

## Exploratory Data Analysis

```{r}
tesla_numeric <- tesla %>%
  select_if(is.numeric)

tesla_model <- tesla_numeric %>%
  select(-more_than_12) %>%
  select(-no_v2) %>%
  select(-no_urban) %>%
  select(-no_v3) %>%
  select(-no_v4) %>%
  select(-total) %>%
  select(-mile)

tesla_model
```

```{r}
tesla_model %>%  # none of our cities contain a v4 charger
  cor() %>%
  corrplot.mixed()
```

## Linear Model

```{r}
linear_model <- lm(no_stations ~ ., data = tesla_model)
summary(linear_model)
```

