---
title: "Tesla Supercharger Analysis"
author: "Jim-Lee Wu and Fion Huang"
subtitle: 
output:
  pdf_document:
    number_sections: true
  html_document:
    number_sections: true
  word_document: default
---
```{r}
library(tidyverse)
library(corrplot)
```


# Introduction

There has been an increasing rise in popularity in electronic vehicles (EVs) throughout the past decade for its environmental, economic, and performance benefits. With so many EVs filling the streets, a need for a proper EV charging infastructure is needed. Just like gasoline cars, EV drivers require charging stations where they can recharge when going long distances. 

## Dataset

```{r}
tesla <- read.csv("Tesla Dataset.csv")
tesla
```

```{r}
tesla %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot.mixed()
```


## How The Data Was Collected

## Parameters

`city` City

`county` County of City

`population` 2022 Population 

`land_size` 2020 Land area in square miles

`medium_income` medium household income in 2018-2022 (in 2022 dollars)

`per_capita_income` per_capita_income: mean income of every person in 2018-2022 (in 2022 dollars)

`no_stations` the number of Tesla Supercharger stations

`mile` the number of stations that are less than one mile from a highway/freeway

`more_than_12` the number of stations that have more than 12 chargers

`no_v2`the number of V2 chargers

`no_urban` the number of urban chargers

`no_v3` the number of V3 chargers

`no_v4` the number of V4 chargers

`total` total number of chargers
